Social Participation Ontology: community documentation, enhancements and use examples

### 2015-10-01 plosOne/plos.tex
repassado resumo com Dilvan do arquivo geral p plos

\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
    Participatory democracy advances in virtually all governments. South America presents a prominent context with mixed culture and social predisposition. In 2012, civil, academic and governmental parties started elaborating the ``Common Vocabulary of Social Participation'' (\vcps\ from the Brazilian name \emph{Vocabul\'ario Comum de Participa\c{c}\~ao Social}), as a public and online process. By May, 2013, first reference documents were publicized, together with a preliminary \owl\ code,
logos, and a diagram of a general ``public consultation''.
The \corais\ platform kept online records of the process, like discussions and preparation of texts. 
This article exposes this material and reports an elementary unfolding: the ``Social Participation Ontology'' (\ops\ from the Brazilian name \emph{Ontologia de Participa\c{c}\~ao Social}). To exhibit this new ontology, these steps were considered: correction of ontological contradictions and \owl\ protocol use errors; completion of \vcps\ \owl\ code into a preliminary version of the \ops; translations and standardizations; enhancements of class names and labels in Portuguese, Spanish and English; a toy expansion of the ontology by further specifying classes; linked data examples regarding dereferencing, a \sparql\ endpoint and participatory instances; use cases by researchers and public managers. Ongoing work involves further adoption of \ops\ by the official Brazilian federal portal of social participation, further adoption by civil participative organizations, and linkage to other participative ontologies. \ops\ is being used as an upper ontology, and all classes linked further to \foaf\ and \bfo\ as higher upper ontologies.

>>>>

\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
    Participatory democracy advances in virtually all governments and especially in South America which presents a mixed culture and social predisposition.
 In 2012, civil, academic and governmental parties started elaborating the ``Social Participation Common Vocabulary'' (\vcps\ from the Brazilian name \emph{Vocabul\'ario Comum de Participa\c{c}\~ao Social}), as a public and online process. By May 2013, first reference documents were publicized, together with preliminary \owl\ code, logos, and a diagram for a general ``public consultation''.
The \corais\ platform kept online records of the process, like discussions and text drafts. 
This article presents this material and proposes, based on it, the ``Social Participation Ontology'' (\ops\ from the Brazilian name \emph{Ontologia de Participa\c{c}\~ao Social}). To create  this new ontology, these steps were followed: correction of ontological contradictions and \owl\ errors in \vcps; inclusion of concepts documented in the \corais\ platform (but not present in \vcps) in a preliminary version of \ops; entity name standardization; translation of labels to Portuguese, Spanish and English; an ontology expansion; creation of linked data examples and  a \sparql\ endpoint; compilation of use cases from researchers and public managers. Ongoing work involves further adoption of \ops\ by the official Brazilian federal portal for social participation and  NGOs, and linkage to other ontologies for participation. The \ops\ is being used as an upper ontology, and all classes linked further to \foaf\ and \bfo\ as higher upper ontologies.
